\modLP{Identification of} Biological Regulatory Networks from Process Hitting models

### 2014-09-12 parts-infer-ig.tex
update

\subsection{Focal Processes}\label{ssec:focal}

\todo{The two following subsections (down to “Interaction Inference”) have a lot of definitions but few examples and not enough informal explanations, especially on their use.
Especially:
\begin{itemize}
  \item What kinds of concept do represent $\focals$?
  \item What is the importance of a strict context?
  \item Add an outline guiding the reader through these subsections.
\end{itemize}}

Many of the inferences defined in the rest of this paper rely on the knowledge of \emph{focal
processes} w.r.t. a given context (a set of processes that are potentially present).
When such a context applies, we expect to (always) reach one focal process in a bounded number of
actions.

For $S_a\subseteq L_a$ and a context (set of processes) $\ctx$, let us define as $\PHa(S_a,\ctx)$
the set of actions on the sort $a$ having their hitter in $\ctx$ and target in $S_a$
(\pref{eq:PHa-ctx});
and the digraph $(V, E)$ where arcs are the bounces within the sort $a$ triggered by actions
in $\PHa(S_a,\ctx)$ (\pref{eq:bounce-graph}).
$\focals(a,S_a,\ctx)$ denotes the set of focal processes of sort $a$ in the scope of
$\PHa(S_a,\ctx)$ (\pref{def:focals}).
\begin{align}
\PHa(S_a,\ctx) & \DEF \{ \PHfrappe{b_i}{a_j}{a_k}\in\PHa \mid b_i\in\ctx \wedge a_j\in S_a \}
\label{eq:PHa-ctx}
\\
\begin{split}
E  & \DEF \{(a_j,a_k)\in (S_a \times \PHl_a) \mid 
			\exists\PHfrappe{b_i}{a_j}{a_k}\in \PHa(S_a,\ctx) \}
\\
V & \DEF S_a \cup \{ a_k\in L_a\mid \exists (a_j,a_k)\in E\}
\end{split}
\label{eq:bounce-graph}
\end{align}

\begin{definition}[$\focals(a,S_a,\ctx)$]\label{def:focals}
The set of processes that are focal for processes in $S_a$ in the scope of $\PHa(S_a,\ctx)$
are given by:
%$\focals(a,S_a,\ctx)$ is the set of focal processes of sort $a$ in the context $\ctx$:
\[
\focals(a,S_a,\ctx) \DEF
\begin{cases}
\{ a_i \in V \mid \nexists (a_i,a_j)\in E\} & \text{if the digraph $(V,E)$ is acyclic},\\
\emptyset & \text{otherwise.}\\
\end{cases}
\]
\end{definition}

We note $\PHl(\ctx)$ the set of states $s\in L$ such that $\forall a\in\PHsort(\ctx), \PHget{s}{a}\in\ctx$,
where $\PHsort(\ctx)$ is the set of sorts with processes in $\ctx$.
We say a sequence of actions $h^1,\dots,h^n$ is \emph{bounce-wise} if and only if
$\forall m\in\segm{1}{n-1}, \PHbounce(h^m)=\PHtarget(h^{m+1})$.
From \pref{def:focals}, it derives that:
\begin{enumerate}
\item if $\focals(a,S_a,\ctx)=\emptyset$, there exists a 
state $s\in \PHl(\ctx\cup S_a)$ such that $\forall n\in\mathbb N$ there
exists a bounce-wise sequence of actions $h^1,\dots,h^{n+1}$ in $\PHa(S_a,\ctx)$ 
with $\PHtarget(h^1)\in s$.
\item if $\focals(a,S_a,\ctx)\neq\emptyset$, for all
state $s\in \PHl(\ctx\cup S_a)$,
for any bounce-wise sequence of actions $h^1,\dots,h^n$ in $\PHa(S_a,\ctx)$ where $\PHtarget(h^1)\in
s$,
either
 $\PHbounce(h^n) \in \focals(a,S_a,\ctx)$,
or
$\exists h^{n+1}\in \PHa(a,\ctx)$ such that $\PHbounce(h^n) = \PHtarget(h^{n+1})$.
Moreover $n\leq|\PHa(S_a,\ctx)|$ (i.e. no cycle of actions possible).
\end{enumerate}

It is worth noticing that those bounce-wise sequences of actions may not be successively playable in
a state $s\in L(\ctx\cup S_a)$.
Indeed, nothing impose that the hitters of actions are present in $s$.
In the general case, the playability of those bounce-wise sequences, referred to as \emph{focals
reachability} may be hard to prove.
However, in the scope of this paper, the particular contexts used with $\focals$ ensure this property.
Notably, the rest of this section uses only \emph{strict} contexts (\pref{def:strict-ctx}) which
allow at most one hitter per sort in the bounce-wise sequences (and thus are present in $s$).

\begin{definition}[Strict context for $S_a$]\label{def:strict-ctx}
A context (set of processes) $\ctx$ is strict for $S_a\subseteq L_a$ if and only if
$\{b_i,b_j\} \subset \ctx \wedge b\neq a \Rightarrow i=j$.
\end{definition}

In other words, assuming focals reachability, if $\focals(a,S_a,\ctx)$ is empty, there exists a
sequence of actions that may be played an unbound number of times (cycle);
if it is non-empty, it is ensured that any state in $\PHl(\ctx\cup S_a)$ converges, in a bounded
number of steps, either to a process in $S_a$ that is not hit by processes in $\ctx$, or to a process in
$L_a\setminus S_a$.

\begin{example}
In the PH of \pref{fig:runningPH-1}, we obtain:
\begin{align*}
\focals(a,L_a,\{b_0,c_0\}) &= \{ a_0 \}
&
\focals(a,L_a,\{b_1,c_1\}) &= \{ a_2 \}
\\
\focals(a,L_a,\{b_1,c_0\}) &= \emptyset
&
\focals(a,\{a_1\},\{b_1,c_0\}) &= \{ a_0, a_2 \}
\end{align*}
\end{example}

@@

@@

Any sort that does not act as a component should then be treated as a cooperative sort.
As explained in \pref{ssec:PH}, the role of a cooperative sort $\upsilon$ is to compute the current
state of set of cooperating processes.
Hence, for each sub-state $\sigma$ formed by the sorts hitting $\upsilon$, $\upsilon$ should
converge to a focal process.
This is expressed by \pref{pro:wf-cooperative-sort}, where
the set of sorts having an action on a given sort $a$ is given by 
$\PHdirectpredec{a}$ (\pref{eq:ph_direct_predec})
and $\PHproc(\sigma)$ is the set of processes that compose the sub-state $\sigma$.

\begin{equation}
\forall a \in \PHs, \PHdirectpredec{a} \DEF \{b \in \PHs \mid \exists \PHfrappe{b_i}{a_j}{a_k}\in\PHa \}
\label{eq:ph_direct_predec}
\end{equation}

\begin{property}[Well-formed cooperative sort]\label{pro:wf-cooperative-sort}
A sort $\upsilon\in\PHs$ is a well-formed cooperative sort if and only if
each configuration $\sigma$ of its predecessors leads $\upsilon$ to a unique focal process,
denoted by $\upsilon(\sigma)$:
\[
\forall \sigma \in {\textstyle\prod_{
a\in\PHdirectpredec{\upsilon} \wedge a\neq \upsilon}}
\PHl_{a},
\focals(\upsilon,\PHl_\upsilon,\PHproc(\sigma)\cup \PHl_\upsilon) = \{ \upsilon(\sigma) \}\]
\end{property}

Such a property allows a large variety of definitions of a cooperative sort, but
for the sake of simplicity, does not allow the existence of multiple focal processes.
While this may be easily extended to (the condition becomes 
$\focals(\upsilon,\PHl_\upsilon, \PHproc(\sigma)\cup \PHl_\upsilon)\neq\emptyset$), it makes some
hereafter equations a bit more complex to read as they should handle a set of focal processes instead
of a unique focal process.

@@

where $\f{sign}(x) = \begin{cases}|x|/x & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$ \enspace.

>>>>

@@

\todo{rework text with new paper structure (\pref{sec:tr2global})}

@@

\todo{maybe we should also support a user-supplied $\Gamma$ and give the
constraints (included in the prev. def?)}

@@

where $\f{sign}(x) = 1$ if $x>0$, $\f{sign}(x) = -1$ if $x < 0$, and
$\f{sign}(x) = 0$ if $x=0$.

### 2014-09-12 parts-ph.tex
update

We recall here the definition and semantics of the Process Hitting (PH), and its usage to model
cooperation between concurrent components.
Two examples of PH modeling a BRN at different abstraction levels are given.
They serve as running examples in the rest of this article.

@@

\end{tikzpicture}
}
\caption{\label{fig:runningPH-1}
A Process Hitting (PH) example.
Sorts are represented by labeled boxes, and processes by circles (ticks are
the identifiers of the processes within the sort, for instance, $a_0$ is the
process ticked $0$ in the box $a$).
An action (for instance $\PHfrappe{b_1}{a_1}{a_2}$) is represented by a pair of
directed arcs, having the hit part ($b_1$ to $a_1$) in plain line and the bounce
part ($a_1$ to $a_2$) in dotted line.
Here, actions involving $b_1$ or $c_1$ are in thick lines.
%The current state is represented by the grayed processes:
%$\state{a_0,b_1,c_0,d_0}$.
}
\end{figure}

@@

While the construction of cooperation in PH allows to encode any boolean functions
between cooperating processes \cite{PMR10-TCSB}, it is worth noticing they introduce a temporal
shift in their application.
This allows the existence of interleaving of actions leading to a cooperative sort representing a
past sub-state of the presence of the cooperative processes.
The resulting behavior is then an over-approximation
of the realization of an instantaneous cooperation.

>>>>

\todo{(LP) rework text, make presentation more CS oriented}

We give the the definition and semantics of the Process Hitting (PH), and its usage to model
cooperation between concurrent components.
Two examples of PH modeling a BRN at different abstraction levels are given.
They serve as running examples in the rest of this article.

@@

\end{tikzpicture}
}
\caption{\label{fig:runningPH-1}
\todo{make more visible}
A Process Hitting (PH) example.
Sorts are represented by labeled boxes, and processes by circles (ticks are
the identifiers of the processes within the sort, for instance, $a_0$ is the
process ticked $0$ in the box $a$).
An action (for instance $\PHfrappe{b_1}{a_1}{a_2}$) is represented by a pair of
directed arcs, having the hit part ($b_1$ to $a_1$) in plain line and the bounce
part ($a_1$ to $a_2$) in dotted line.
Here, actions involving $b_1$ or $c_1$ are in thick lines.
%The current state is represented by the grayed processes:
%$\state{a_0,b_1,c_0,d_0}$.
}
\end{figure}

@@

### 2014-09-12 parts-tr2global.tex
update

% vim:spell spelllang=en:
\section{Interaction Graph Inference from Process Hitting}\label{sec:infer-IG}

The Interaction Graph (IG) is an abstract representation of the direct qualitative influences,
positive and/or negative, between the components of the system.
As discussed in \pref{sec:intro}, the IG allows to efficiently characterize global dynamical
properties for the concrete system, such as the capability for multi-stationarity or oscillation.

In a typical biological network modeling process, a prior IG is generally the starting point for
the formal system specification.
However, it is common that the prior IG actually refers to interactions that reveal to be 
non-effective with respect to the dynamics.
Hence, deriving the IG directly from the dynamical models lead to more concise IGs, enhancing the
conclusiveness of static analyses based upon this abstract representation.

In this section, we formally derive the IG corresponding to a given PH that is well-formed for BRN
modeling.
This section first introduces the notion of focal processes within a PH (\pref{ssec:focal})
which is used to characterize well-formed PH for IG inference (\pref{ssec:wf}).
%and as well used by the parametrization inference presented in \pref{sec:infer-K}.
The rules for inferring the interactions between components from a PH are
described in \pref{ssec:infer-IG}.
We consider hereafter a global PH $(\PHs,\PHl,\PHa)$ on which the IG inference is to be
performed.

@@

Many of the inferences defined in the rest of this paper rely on the knowledge of \emph{focal
processes} w.r.t. a given context (a set of processes that are potentially present).
When such a context applies, we expect to (always) reach one focal process in a bounded number of
actions.

@@

\subsection{Well-formed Process Hitting for Interaction Graph Inference}\label{ssec:wf}

The inference of an IG from a PH assumes that the PH defines two types of sorts:
the sorts corresponding to BRN components, and the cooperative sorts.
This leads to the characterization of a \emph{well-formed} PH for IG inference.

The identification of sorts modeling components relies on the observation that their processes
represent (ordered) qualitative levels.
Hence an action on such a sort cannot make it bounce to a process at a distance more than one.
The set of sorts satisfying such a condition is referred to as $\Gamma$
(\pref{eq:PH-components}).
Therefore, in the rest of this paper, $\Gamma$ denotes the set of components of the BRN to infer.

\begin{equation}
\Gamma \DEF \{a \in \PHs \mid \nexists \PHfrappe{b_i}{a_j}{a_k} \in \PHa, |j - k| > 1\} \\
\label{eq:PH-components}
\end{equation}

@@

Finally, \pref{pro:wf-ph} sums up the conditions for a Process Hitting to be suitable for IG
inference.
In addition of having either component sorts or well-formed cooperative sorts, we also require that
there is no cycle between cooperative sorts, and that
sorts being never hit (\ie{} serving as an invariant environment) are components.

\begin{property}[Well-formed Process Hitting for IG inference]\label{pro:wf-ph}
A PH is well-formed for IG inference if and only if the following conditions are verified:
\begin{itemize}
\item 
each sort $a\in\PHs$ either belongs to $\Gamma$, or is a well-formed cooperative sort;
\item 
there is no cycle between cooperative sorts
(the digraph $(\Sigma,\{(a,b)\in(\Sigma\times\Sigma)\mid \exists \PHfrappe{a_i}{b_j}{b_k}\in\PHa
\wedge a\neq b\wedge \{a,b\}\cap\Gamma=\emptyset \})$ is
acyclic);
\item 
sorts having no action hitting them belong to $\Gamma$
($\{ a \in \Sigma\mid \nexists \PHfrappe{b_i}{a_j}{a_k}\in\PHa\} \subset \Gamma$).
\end{itemize}
\end{property}

\begin{example}
In the PH of \pref{fig:runningPH-2}, $bc$ is a well-formed cooperative sort as defined in \pref{pro:wf-cooperative-sort}, because:
\begin{align*}
\focals(bc, \PHl_{bc}, \{b_0, c_0\} \cup \PHl_{bc}) = \{bc_{00}\} && \focals(bc, \PHl_{bc}, \{b_0, c_1\} \cup \PHl_{bc}) = \{bc_{01}\} \\
\focals(bc, \PHl_{bc}, \{b_1, c_0\} \cup \PHl_{bc}) = \{bc_{10}\} && \focals(bc, \PHl_{bc}, \{b_1, c_1\} \cup \PHl_{bc}) = \{bc_{11}\}
\end{align*}
Hence, both \pref{fig:runningPH-1} and \pref{fig:runningPH-2} are well-formed PH for IG inference
with $\Gamma = \{a,b,c\}$.
\end{example}

\subsection{Interaction Inference}\label{ssec:infer-IG}

At this point we can divide the set of sorts $\PHs$ into components ($\Gamma$, see \pref{eq:PH-components}) and cooperative sorts
($\PHs \setminus \Gamma$) that will not appear in the IG. 
We define in \pref{eq:ph_predec} the set of predecessors of a sort $a$, that is, the sorts influencing $a$
by considering direct actions and possible intermediate cooperative sorts.
The predecessors of $a$ that are components are the regulators of $a$, denoted $\PHpredecgene{a}$
(\pref{eq:regulators}).
\begin{align}
\begin{split}
\forall a \in \PHs, \PHpredec{a} &\DEF \{b \in \PHs \mid \exists n \in \mathbb{N}^*, \exists
(c^k)_{k \in \segm{0}{n}} \in \PHs^{n+1}, \\
                                   & \quad \quad c^0 = b \wedge c^n = a \\
                                   & \quad \quad \wedge \forall k \in \segm{0}{n-1},
                   c^k \in \PHdirectpredec{c^{k+1}} \cap (\PHs\setminus\Gamma)\}
\end{split}
\label{eq:ph_predec}
\\
\forall a\in \PHs, \PHpredecgene{a} & \DEF \PHpredec{a} \cap \Gamma
\label{eq:regulators}
\end{align}

Given a set $g$ of components and a configuration (\ie a sub-state) $\sigma$, $\ctx_g(\sigma)$
refers to the set of processes hitting $a$ regulated by any sort in $g$ (\pref{eq:ctx-sigma}).
If $g=\{b\}$, we simple note $\ctx_b(\sigma)$.
This set is composed of the active processes of sorts in $g$, and the focal process (assumed
unique) of the cooperative sorts $\upsilon$ hitting $a$ that have a predecessor in $g$.
The evaluation of the focal process of $\upsilon$ in context $\sigma$, denoted $\upsilon(\sigma)$,
relies on \pref{pro:wf-cooperative-sort}, which gives its value when all the direct predecessors of
$\upsilon$ are defined in $\sigma$.
When a predecessor $\upsilon'$ is not in $\sigma$, we extend the evaluation by recursively computing
the focal value of $\upsilon'$ is $\sigma$, as stated in \pref{eq:cooperative-eval},
where $\uplus$ stands for a union of states (which are sets of processes).
Because there is no cycle between cooperative sorts, this recursive evaluation of $\upsilon(\sigma)$
always terminates.
\begin{align}
\forall g\subset \Gamma,
  \ctx_g(\sigma) & \DEF \{ \sigma[b] \mid b\in g \} \cup \{ \upsilon(\sigma) \mid
\upsilon\in\PHdirectpredec{a} \setminus \Gamma \wedge g\cap \PHpredecgene{\upsilon} \neq \emptyset \}
\label{eq:ctx-sigma}
\\
\upsilon(\sigma) & \DEF
\upsilon(\sigma \uplus \state{\upsilon'(\sigma) \mid 
  \upsilon'\in\PHdirectpredec{\upsilon} \wedge
  \upsilon'\in\PHs\setminus\Gamma })
\label{eq:cooperative-eval}
\end{align}

This inference mainly focuses on the presence of actions betweens two sorts to conclude on the presence of a regulation.
We aim at inferring that $b$ activates (inhibits) $a$ if there exists a configuration where increasing
the level of $b$ makes possible the increase (decrease) of the level of $a$,
which is directly inspired from the works of~\cite{Richard2010378}.
To do so, the sets of components cooperating together to hit $a$, called groups of influence of $a$, are studied.
Such groups are given by $X(a)$ which is the set of connected components in the graph linking two regulators
$b$ and $c$ of $a$ if they use a common cooperative sort to have an influence on $a$ (\pref{eq:influence-groups}).
\begin{equation}
X(a) = \mathcal C\left( (\PHpredecgene{a}, \{ \{b,c\} \mid
        \exists \upsilon\in \PHdirectpredec{a} \setminus \Gamma,
        \{b,c\} \subset \PHpredecgene{\upsilon} \}) \right)
\label{eq:influence-groups}
\end{equation}

To infer influence of a component $b$ on a component $a$,
we observe the direction of evolution of $a$ for different active processes of $b$.
The fact that $a$ tends to evolve differently when $b$ changes denotes an influence from $b$.
Formally, for a given group $g$ of regulators of a component $a$
(that is, a minimal set of components regulating $a$ through common cooperative sorts),
and a configuration $\ctx$ on $g$, we note
$\irB_a(\sigma)$ the set of processes towards which $a$ can bounce (\pref{eq:possible-bounces}).
%This set is defined using the set $\irF_a(\sigma)$ of action hitting $\PHget{\sigma}{a}$ in $\sigma$ (\pref{eq:possible-actions})
%and the set of processes $\ctx$ hitting any process of $a$ (\pref{eq:ctx-sigma}).
If $a$ cannot be hit by any action in $\sigma$, then $\irB_a(\sigma) = \{ \PHget{\sigma}{a} \}$.
%
\begin{align}
\begin{split}\label{eq:possible-bounces}
  &\forall g \in X(a), \forall \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c,
  \irB_a(\sigma) \DEF 
  \begin{cases}
    \irF_a(\sigma)
      & \text{ if } \irF_a(\sigma) \neq \emptyset\\
    \{ \PHget{\sigma}{a} \}
      & \text{ if } \irF_a(\sigma) = \emptyset
  \end{cases}\\
  &\text{where: } \irF_a(\sigma) \DEF \{ a_k \in \PHl_a \mid \exists b \in \PHs, \exists \PHfrappe{b_i}{a_j}{a_k} \in \PHa,\\
  & \qquad\qquad\qquad\qquad\qquad\qquad \PHget{\sigma}{a} = a_j \wedge \PHget{(\ctx_{g}(\sigma))}{b} = b_i \}\\
\end{split}
\end{align}

\pref{pps:inference-edges} details the inference of all existing influences between components occurring
with a threshold $t$.
The main idea of this inference is that
the presence of a positive (negative) influence of a component $b$ on $a$ denotes the fact that
there exists a state in which increasing the level of $b$ tends to make the future level of $a$ rise (drop)
(\pref{eq:edges-inference}).
Therefore, these influences are split into positive and negative ones, and represent possible edges in the final IG.
Furthermore, studying the influences of the groups of regulators of $a$
allows to study its auto-influences, and thus infer auto-edges on $a$ in the IG (\pref{eq:edges-inference-auto}).
Finally, \pref{eq:edges-inference-noreg} handles the special case where $a$ has no regulators.
We ignore the cases where a component has no visible influence on another.
%
\begin{proposition}[Influences inference]\label{pps:inference-edges}
We define the set of positive (resp. negative) influences $\hat{E}_+$ (resp. $\hat{E}_-$) for any $a\in\Gamma$ by:
\begin{align}
% Arcs a -> b, a ≠ b
\begin{split}\label{eq:edges-inference}
  \forall b\in\PHpredecgene{a}, \forall s \in \{ +, - \}, \\
  b \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), b \in g,
  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
    &\qquad s = \f{sign}(k - a)
\end{split}\\
% Auto-arcs
\begin{split}\label{eq:edges-inference-auto}
  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \exists g \in X(a), \exists b \in g,
  \exists \sigma \in \textstyle\prod_{c\in g \cup \{ a \}} \PHl_c, \\
    &\qquad \{ b_t, b_{t+1} \} \subset \PHl_b \wedge b_t \in \sigma,\\
    &\qquad \exists a_j \in \irB_a(\sigma), \exists a_k \in \irB_a(\sigma\{b_{t+1}\}), \\
    &\qquad s = \f{sign}(a_k - a_j)
\end{split}\\
\begin{split}\label{eq:edges-inference-noreg}
  \forall s \in \{ +, - \}, \quad\qquad\qquad \\
  a \xrightarrow{t+1} a \in \hat{E}_s \Longleftrightarrow\ & \reg{a} = \emptyset \wedge \{ a_t, a_{t+1} \} \subset \PHl_a, \\
    &\qquad \exists a_j \in \irB_a(\PHetat{a_t}), \exists a_k \in \irB_a(\PHetat{a_{t+1}}), \\
    &\qquad s = \f{sign}(a_k - a_j)
\end{split}
\end{align}
where $\f{sign}(x) = \begin{cases}|x|/x & \text{ if $x \neq 0$} \\ 0 & \text{ if $x = 0$}\end{cases}$.
\end{proposition}

We are now able to infer the edges of the final IG by considering positive and negative influences
(\pref{pps:inference-IG}).
We infer a positive (resp. negative) edge if there only exist corresponding influences with the same sign.
If an influence is both positive and negative, we infer an unsigned edge.
In the end, the threshold of each edge is the minimum threshold for which an influence has been found.
%
\begin{proposition}[Interaction Graph inference]\label{pps:inference-IG}
We infer $\IG = (\Gamma,E)$ using \pref{pps:inference-edges} as follows:
\begin{align*}
E_+ &= \{ \GRNedge{a}{+}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_-
  \wedge t = \min \{ l \mid a \xrightarrow{l} b \in \hat{E}_+\}\} \\
E_- &= \{ \GRNedge{a}{-}{t}{b} \mid \nexists a \xrightarrow{t'} b \in \hat{E}_+
  \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_-\}\} \\
E_\pm &= \{ \GRNedge{a}{\uns}{t}{b} \mid \exists a \xrightarrow{t'} b \in \hat{E}_+ \wedge \exists a \xrightarrow{t''} b \in \hat{E}_- \\
  & \qquad\qquad\qquad \wedge t = \min \{l \mid a \xrightarrow{l} b \in \hat{E}_- \cup \hat{E}_+\}\}
\end{align*}
\end{proposition}

\begin{example}
The IG inference of the PH of \pref{fig:runningPH-1} gives the
IG in \pref{fig:BRN-inf1}, containing the following edges:
\begin{align*}
  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}, \GRNedgef{a}{+}{1}{a}, \GRNedgef{b}{+}{1}{b}, \GRNedgef{c}{+}{1}{c}\}\\
  E_- &= \{\GRNedgef{a}{-}{2}{b}\} \qquad\qquad\qquad\qquad\qquad
  E_\uns = \emptyset
\end{align*}
This IG is close to the one in \pref{fig:runningBRN} but not equivalent,
as each component has an additional auto-action.
The auto-actions on $b$ and $c$ are the consequence of a global stability
in some configurations: $c$ never evolves, and neither does $b$ when $a_2$ is not active.
The auto-action on $a$ is mainly caused by its multi-valued nature.

The inference of the refined PH of \pref{fig:runningPH-2} gives the same IG.

\begin{figure}[t]
\centering
\scalebox{1.2}{
\begin{tikzpicture}[grn]
  \path[use as bounding box] (-1.3,-0.75) rectangle (3.5,1.5);
  \node[inner sep=0] (a) at (2,0) {a};
  \node[inner sep=0] (b) at (0,0) {b};
  \node[inner sep=0] (c) at (2,1.2) {c};
  \path[->]
    (b) edge[bend right] node[elabel, below=-2pt] {$+1$} (a)
    (c) edge node[elabel, right=-2pt] {$+1$} (a)
    (a) edge[bend right] node[elabel, above=-5pt] {$-2$} (b)
    (b) edge[in=-15+180, out=15+180, loop] node[elabel, left=-2pt] {+1} (b)
    (c) edge[in=15, out=-15, loop] node[elabel, right=-2pt] {+1} (c)
    (a) edge[in=15, out=-15, loop] node[elabel, right=-2pt] {+1} (a);
\end{tikzpicture}
}
\caption{\label{fig:BRN-inf1}
  Result of the IG inference performed on the PH of \pref{fig:runningPH-2}.
}
\end{figure}

\end{example}

\begin{example}
If we add the action $\PHfrappe{a_2}{b_0}{b_1}$ to the PH of \pref{fig:runningPH-2},
then two unsigned edges towards $b$ are inferred instead of the previous signed edges:
\begin{align*}
  E_+ &= \{\GRNedgef{b}{+}{1}{a}, \GRNedgef{c}{+}{1}{a}, \GRNedgef{a}{+}{1}{a}, \GRNedgef{c}{+}{1}{c}\}\\
  E_- &= \emptyset \qquad\qquad\qquad\qquad
  E_\uns = \{\GRNedgef{a}{\uns}{2}{b}, \GRNedgef{b}{\uns}{1}{b}\}
\end{align*}
This is due to the fact that the actions $\PHfrappe{a_2}{b_1}{b_0}$ and $\PHfrappe{a_2}{b_0}{b_1}$
introduce an oscillation only caused by $a$, which cannot be represented in Thomas modeling.
\end{example}

>>>>

% vim:spell spelllang=en:
\section{From Local Transitions to Global Functions}\label{sec:tr2global}

\todo{(LP) motivation, outline}
Many of the inferences defined in the rest of this paper rely on the knowledge of \emph{focal
processes} w.r.t. a given context (a set of processes that are potentially present).
When such a context applies, we expect to (always) reach one focal process in a bounded number of
actions.

@@

\todo{(LP) more explanations, check strict context importance, more examples?}

@@

\subsection{Restrictions for Determinism} % or Well-formed Cooperative Sorts

\todo{(LP) redo text}

@@
